EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-07 draft.tex
Changing Images

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We only tested 0, 1 and .5, there is not really a justification - M 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

% What is the main objective of this experiment? Measure scalability?
% - JJ 
% Yes, see below - M

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.

We executed two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list.

with results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

@@

\subsection{Experimental Results} 
\label{results}
% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
% Removed - M
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%

@@

\begin{table}[h!tbp]
    \small
    \caption{Median evaluation rate per second, for each number of workers and dimension}
        % mark significant differences in some way - JJ

@@

\begin{figure}[h!p]
    \centering
    \includegraphics[width=\textwidth]{evalsworker}
    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
    % Which function? - JJ
    \label{fig:spworker}
\end{figure}

>>>>

The experiments can be easily reproduced by using the configuration
files the the EvoSwarm application uses to run the noiseless BBOB testbed
experiments as seen in Section~\ref{sec:exeperimentalDefinition}. % Specify this a bit more (with commands such as
             % docker-compose up and so on and use an itemize
             % environment to highlight it and answer to the reviewer
             % - JJ
As a first step, we must deploy the docker application using a
{\tt docker-compose.yml} file, where the number of worker containers is
specified. Then a JSON file containing the configuration parameters of the
experiment has to be provided.  Table \ref{tab:params} shows an example of these
parameters. The {\em GA-PSO Ratio} parameter indicates the proportion of
populations that will use the GA algorithm.  In the example, with a value of
$0.50$ there will be about the same proportion of GA and PSO populations. If we
specify a value of $0$, this will give us an algorithm with only PSO
populations, and finally, a value of  $1$ means that all populations will run
the GA algorithm. % Opening up the door to ask for testing with
                  % different proportion. You have to justify this
                  % value - JJ  
                  % We mention earlier that we replicated the params of another paper - M 
Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

% What is the main objective of this experiment? Measure scalability?
% - JJ 
% Yes, see below - M
\subsection{Setup}

To evaluate the effect of these parameters on the scalability of the system, we
propose  an experiment in which we selected $f_4$ (Skew Rastrigin-Bueche separable) from the
BBOB testbed. This function has been used because it is computationally demanding,
and in higher dimensions has been difficult for PSO \cite{el2009black} and GAs
\cite{nicolau2009application} to solve with the required FEs. As we
are only comparing in terms of execution time, the results from a single function
can be better understood, since there are fewer factors involved.
Following the procedure described in the previous Section, we executed 
two sets of experiments with five and ten initial populations,
repeating each experiment 30 times using  1, 2, 4, and 8 workers.  We kept the same 
maximum number of FEs, changing the relevant parameters for this. 
See Tables \ref{tab:params:10} and \ref{tab:params:5} 
for the complete list.

\begin{figure}[h!p]
  \centering
  \includegraphics[width=\textwidth]{time_wrkr}
  \caption{time required for the execution of 15 function
  instances, in seconds }

  % Which function? - JJ
  \label{fig:spworker}
\end{figure}

\begin{figure}[h!p]
  \centering
  \includegraphics[width=\textwidth]{evalsworker}
  \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }

  % Which function? - JJ
  \label{fig:spworker}
\end{figure}

\subsection{Experimental Results} 
\label{results}

With results summarized in Figure \ref{fig:spworker}.  The
experiment consisted of executing the same benchmark on EvoSwarm twice but
changing the number of populations used in each run. This is mainly
intended as a showcase of the flexibility of the tool presented in
this paper, but we also wanted to test how the number of workers
affected the time needed to reach a solution. Adding more populations
might add more overhead, so this might have a noticeable effect.
The left column in that figure has the
results when using five populations, and the right side shows the results for
ten. In principle the main issue is how to find the right number of
workers for a particular problem, but in general we wanted to check
the communication overhead when the size of the problem increases and
how the presented framework reacted to it. % Added some explanation
% here - JJ

@@

% that's pretty obvious. Say what you intend to prove with the experiments and why - JJ 
% Removed - M
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%

@@

\begin{table}[h!tbp]
    \small
    \caption{Boxplot of the total number of evaluations for 30 instances of function $f_4$ , with different number of workers and populations sizes }
      % mark significant differences in some way - JJ

@@
